---
title: "York University - Cyclical Program Review - 2014"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,  fig.align='center')

```
# Undergraduate Program in Inforamtion Technology

**Student Questionnaire Results **

*Recreation of the report using Rmarkdown*

*By: Chethana Wickramasinghe (1/31/2020)*


```{r ,include=FALSE}

library(tidyverse)
library(readr)
library(openxlsx)
library(stringi)
library(ggplot2)
library(scales)
library(dplyr)
library(knitr)
library(kableExtra)
library(sqldf)

```

```{r files,include=FALSE}

#read the file with paper submission
#rf: cleaned data extracted from the excel file
#########################################################CHANGE WHEN REQUIRED###########
rf2 <- read.xlsx("./2014_Cyclical_Program_Review_Data.xlsx") 

#remove @sign from the headings
names(rf2) <- gsub("@", "_", names(rf2), fixed = TRUE)

#filter Information technology students
IT_rf2 <-  rf2 %>% filter (program_name == 'Information Technology')
#########################################################
```

```{r functions,include=FALSE, echo=FALSE, warning=FALSE}
#functions
extract_data <- function(sub_df1,array,number){
  col_names <- colnames(sub_df1)

  col_index = 0
  
  for (i in col_names) {
  
  c9_2  <- sub_df1 %>%
    group_by(sub_df1[[i]]) %>%
    summarise(Count = n())
  
  col_index =col_index +1
  colnames(c9_2)[1] <- col_names[col_index]
  colnames(c9_2)[2] <- array[col_index ]
  
  nc9_2 <- c9_2
  colnames(nc9_2)[1] <- "number"
  statement = paste ("SELECT *
                FROM number
                LEFT JOIN nc9_2 USING(number)")
  
  finaldata <- sqldf(statement)
  number <- finaldata
  }
  
  clean<- finaldata %>% select(number, starts_with("Count"))
  clean<- t(clean)
  clean<- as.data.frame(clean)
  clean<- clean[-1,]

  
  return(clean)
}

get_summary <- function(sub_df1){
  na <- sum(length(which(is.na(sub_df1[[1]]))))
  if (na > 0) {
    Frequency_extract <- sub_df1 %>%
    group_by(sub_df1[[1]]) %>%
    summarise(Frequency = n()) %>%
    slice(-n())
  }else{
    Frequency_extract <- sub_df1 %>%
    group_by(sub_df1[[1]]) %>%
    summarise(Frequency = n())
  }
  
  Percent_extract <- sub_df1 %>%
    group_by(sub_df1[[1]]) %>%
    summarise(Percent = n()/nrow(sub_df1) *100)
  
  results<-left_join(Frequency_extract, Percent_extract)
  
  results$Valid_Percent= 100*(results$Frequency)/(sum(results$Frequency))
  
  results$cumulative_Percent= 100*cumsum(results$Frequency)/sum(results$Frequency)

  return(results)
  
}
```

```{r p1 , include=TRUE, echo=FALSE, warning=FALSE}
g1_d <- rf2 %>%
  group_by(program_name) %>%
  summarise(Count = n()) %>%
  mutate( program_name = factor(program_name),
    Ratio = Count / sum(Count),
         label = percent(Ratio ))%>%drop_na()

g1 <-ggplot(g1_d, aes(x=program_name,y=Ratio,label=label,fill=program_name)) +
  geom_bar(stat='identity', fill="#92BAD3") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3), plot.title = element_text(hjust = 0.5))+geom_text(vjust=0.1)+labs(title = "Response Rates By Program", x = "Programs", y ="Response Rate")

g1

```

# Overall program rating

```{r c7 , include=TRUE, echo=FALSE, warning=FALSE}
sub_df1 <- IT_rf2 %>% select(7)
results <- get_summary(sub_df1)

#rename rownames
results$`sub_df1[[1]]` <- c("Excellent","Very good","Good","Fair","Poor","Very poor")
colnames(results)[1] <- " "

results %>%
  kable() %>%
  kable_styling()

```


```{r p2 , include=TRUE, echo=FALSE , warning = FALSE}

fa7 <- IT_rf2 %>% filter(a7 != 7)

#summarising
g2_d <- fa7 %>%
  group_by(a7) %>%
  summarise(Count = n()) %>%
  mutate( a7 = factor(a7),
    Ratio = Count / sum(Count),
         Percentage = round((Ratio *100 ),2))%>%drop_na()#########

#creating the graph
g2 <-ggplot(g2_d, aes(x=a7,y=Ratio,label=Percentage,fill=a7)) +
  geom_bar(stat='identity', fill="#92BAD3") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.3), plot.title = element_text(hjust = 0.5))+geom_text(vjust=1)+labs(title = "Overall program rating",x="Ratings",y ="Percent")+scale_x_discrete(labels=c("Excellent","Very good","Good","Fair","Poor","Very poor"))

g2



```

# Courses achieved these goals

```{r b1 , include=TRUE, echo=FALSE , warning = FALSE}
#extracting the columns incuded only for this question

sub_df1 <-  IT_rf2 %>% select(10:20) 
array <- c("Count1a", "Count2b", "Count3c", "Count4d", "Count1e"
           , "Count2f", "Count3g", "Count4h", "Count1i", "Count2j"
           , "Count3k")
number <- data.frame(number = c("1","2","3","4","5","6","7"))

clean <- extract_data(sub_df1, array, number)
#make first row the column and rename columns
colnames(clean) <- c("All", "Most", "About half", "A few", "None", "No opinion", "Does not apply")

#rename rownames
rownames(clean) <- c("Course instructors provide high quality teaching",
                        "Instructors are enthusiastic about their teaching",
                        "Instructors encourage students to become independent learners",
                        "In 3000- and 4000 - level courses, instructors integrate research into their teaching",
                        "Instructors provide complete course outlines that clarify the course objectives and requirements",
                        "Teaching assistants provide high quality teaching",
                        "Teaching assistants provide interesting additional material to enhance courses",
                        "Lab assistants/technicians make sure that equipment is ready & in working order & supplies ready for
labs",
                        "The balance between lectures, tutorials, labs and other methods makes for effective learning",
                        "Course work and examinations are evaluated fairly",
                        "Every student is treated fairly and equitably")

clean  %>%
  kable() %>%
  kable_styling


```

# Program contributed to:

```{r b2 , include=TRUE, echo=FALSE , warning = FALSE}
#extracting the columns incuded only for this question

sub_df1 <-  IT_rf2 %>% select(21:38) 
array <- c("Count1a", "Count2b", "Count3c", "Count4d", "Count1e"
           , "Count2f", "Count3g", "Count4h", "Count5i", "Count1j"
           , "Count2k", "Count3l", "Count4m" , "Count5n", "Count1o"
           , "Count2p" , "Count3q", "Count4r")

number <- data.frame(number = c("1","2","3","4","5"))

clean <- extract_data(sub_df1, array, number)

#make first row the column and rename columns
colnames(clean) <- c("Very Much", "Quite a bit", "Some", "Very little","Does not apply")

#rename rownames
rownames(clean) <- c("Understanding theories and concepts",
                        "Understanding the limits of your own knowledge",
                        "Understanding different disciplinary approaches to problem solving",
                        "Understanding different methodological approaches",
                        "Understanding major ethical guidelines",
                        "Critically evaluate arguments in analyzing conflicting views and develop a standpoint",
                        "Ability to reason with numbers",
                        "Ability to develop and evaluate creative solutions",
                        "Understanding professional codes of practice",
                        "Identify and address your own learning needs",
                        "Critically evaluate communications",
                        "Writing effectively",
                        "Present ideas effectively",
                        "Exercising initiative and accepting responsibility",
                        "Working effectively in a team",
                        "Leadership ability",
                        "Respect for, and support of, diversity",
                        "Developing a concern for the well-being of others")


clean  %>%
  kable() %>%
  kable_styling

```

# Contribution to knowledge, etc.

```{r b3 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(39:42)
array <- c("Count1", "Count2", "Count3", "Count4")
number <- data.frame(number = c("1","2","3","4","5","6"))
clean <- extract_data(sub_df1, array, number)

#make first row the column and rename columns
colnames(clean) <- c("Very Much", "Quite a bit", "Some", "Very little","Cannot say", "Not applicable")

#rename rownames
rownames(clean) <- c("Required courses", "Foundation courses", "
                        General education courses","Elective courses")

clean  %>%
  kable() %>%
  kable_styling


```

# Program

```{r c1 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(43:45)
array <- c("Count1", "Count2", "Count3")
number <- data.frame(number = c("1","2","3","4","5"))
clean <- extract_data(sub_df1, array, number)

#make first row the column and rename columns
colnames(clean) <- c("Very well", "Fairly well", "Not very well", "Not at all well","Cannot say")

#rename rownames
rownames(clean) <- c("Covered a variety of disciplinary or multi-disciplinary
perspectives ", "Addressed equity, diversity, inclusivity", "Avoided unnecessary repetition of material in other
courses")

clean  %>%
  kable() %>%
  kable_styling

```

# Contribution to preparation for upper year courses

```{r c3 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(46:49)
array <- c("Count1", "Count2", "Count3", "Count4")
number <- data.frame(number = c("1","2","3","4","5","6"))

clean <- extract_data(sub_df1, array, number)

#make first row the column and rename columns
colnames(clean) <- c("Very Much", "Quite a bit", "Some", "Very little","Cannot say", "Not applicable")

#rename rownames
rownames(clean) <- c("Required courses", "Foundation courses", "
                        General education courses","Elective courses")

clean  %>%
  kable() %>%
  kable_styling

```

# Difficulty enrolling

```{r c4 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(50:51)
array <- c("Count1", "Count2")
number <- data.frame(number = c("1","2","3","4","5"))
clean <- extract_data(sub_df1, array, number)

#make first row the column and rename columns
colnames(clean) <- c("Always","Most of the time", "Sometimes", "Rarely","Never")

#rename rownames
rownames(clean) <- c("The course was full","The course was not offered at a time that was convenient for you" )

clean  %>%
  kable() %>%
  kable_styling

```

# Courses preferred

```{r c5 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(52:54)
array <- c("Count1", "Count2", "Count3")
number <- data.frame(number = c("1","2","3","4","5","6"))
clean <- extract_data(sub_df1, array, number)

#make first row the column and rename columns
colnames(clean) <- c("All", "Most", "About half", "A few", "None" , "Not sure")

#rename rownames
rownames(clean) <- c("8:30 am - 4:30 pm", "4:30 pm - 6:30 pm","7:00 pm - 10:00 pm")

clean  %>%
  kable() %>%
  kable_styling

```

# Courses

```{r c6 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(55:57)
array <- c("Count1", "Count2", "Count3")
number <- data.frame(number = c("1","2","3","4","5","6"))
clean <- extract_data(sub_df1, array, number)

#make first row the column and rename columns
colnames(clean) <- c("All", "Most", "About half", "A few", "None" , "Not sure")

#rename rownames
rownames(clean) <- c("Used technology to enhance
learning experience ", "Provided hands-on learning
opportunities","Like to take online")

clean  %>%
  kable() %>%
  kable_styling

```

# Times contacted

```{r c14 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(61:65)
array <- c("Count1", "Count2", "Count3", "Count4", "Count5")
number <- data.frame(number = c("1","2","3","4"))
clean <- extract_data(sub_df1, array, number)

#make first row the column and rename columns
colnames(clean) <- c("Never", "Once", "Twice", "Three or more times")

#rename rownames
rownames(clean) <- c("Professor in program", "Undergraduate Program Director ","Undergraduate program office" ,"Student Academic Services","Registrar's office")

clean  %>%
  kable() %>%
  kable_styling

```

# Extent received when needed

```{r d2 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(66:69)
array <- c("Count1", "Count2", "Count3", "Count4", "Count5")
number <- data.frame(number = c("1","2","3","4","5"))
clean <- extract_data(sub_df1, array, number)

#make first row the column and rename columns
colnames(clean) <- c("Very much","Quite a bit","Some", "Very little", "Does not apply")

#rename rownames
rownames(clean) <- c("Academic advising", "Information about important deadlines","Assistance understanding academic rules and
policies" ,"Information about academic support options")

clean  %>%
  kable() %>%
  kable_styling

```

# How likely to enroll in these programs: 

```{r c9_c10 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(58:59)
array <- c("Count1", "Count2")
number <- data.frame(number = c("1","2","3","4","5","6"))
clean <- extract_data(sub_df1, array, number)

#make first row the column and rename columns
colnames(clean) <- c("Very likely","Somewhat likely", "Not very likely","Not at all likely", "Not sure at this time","Already enrolled")

#rename rownames
rownames(clean) <- c("Internship", "Cooperative Education")

clean  %>%
  kable() %>%
  kable_styling

```

# Other 'hands-on' learning opportunities 


```{r c11 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(60)

results <- get_summary(sub_df1)

#rename rownames
results$`sub_df1[[1]]` <- c("Very interested","Somewhat interested", "Not very interested","Not at all interested", "Not sure at this time")

colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling

```


# Quality of academic advising 

```{r c16 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(97)
results <- get_summary(sub_df1)

results$`sub_df1[[1]]` <- c("Excellent","Very Good", "Good","Fair","Poor","Very Poor","Never Obtained Advice")

colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling

```

# Quality of program office service 

```{r c18 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(98)
results <- get_summary(sub_df1)

results$`sub_df1[[1]]` <- c("Excellent","Very Good", "Good","Fair","Poor","Very Poor")
colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling

```

# Quality of library resources

```{r c19 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(99)

results <- get_summary(sub_df1)

results$`sub_df1[[1]]` <- c("Excellent","Very Good", "Good","Fair","Poor","Very Poor","Not Sure")

colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling

```

# Quality of service of: 

```{r d6 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(70:74)
array <- c("Count1", "Count2","Count3", "Count4","Count5")
number <- data.frame(number = c("1","2","3","4","5","6","7","8"))
clean <- extract_data(sub_df1, array, number)

#make first row the column and rename columns
colnames(clean) <- c("Excellent","very Good","Good","Fair","Poor","Very Poor","Never used","Never heard of service")

#rename rownames
rownames(clean) <- c("York University English
Language Institute", "Writing Support Centre","Learning Skills Services","Learning Commons","Counselling and Disability
Services")

clean  %>%
  kable() %>%
  kable_styling

#I give a frequency to each word of this list 
b <- sample(seq(0,1,0.01) , length(colnames(clean)) , replace=TRUE) 
 
#The package will automatically make the wordcloud ! (I add a black background)
par(bg="black") 
#wordcloud(colnames(clean) , b , col=terrain.colors(length(colnames(clean)) , alpha=0.9) , rot.per=0.3 )

```

# Disability or continuing health condition that requires accommodation

```{r d7 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(75)

results <- get_summary(sub_df1)

results$`sub_df1[[1]]` <- c("Yes","No")

colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling

```

# Need for accommodation generally met

```{r d25 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(76)

results <- get_summary(sub_df1)

results$`sub_df1[[1]]` <- c("Fully","Partly","Not met")
colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling

```

# Paid job during September 2013-April 2014 academic year

```{r f1 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(78)

results <- get_summary(sub_df1)

results$`sub_df1[[1]]` <- c("No","Yes")

colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling

```

# Paid work

```{r f2 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(79)

no_na <- sub_df1 %>% na.omit

results <- data.frame("Paid_work" = "Average hours per week", "N" = nrow(no_na), "Minimum" = min(no_na), "Maximum" = max(no_na) ,"Mean" = mean(no_na$f2))


results  %>%
  kable() %>%
  kable_styling

```

# Due to job, spent less time on campus than would like

```{r f4 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(80)

results <- get_summary(sub_df1)

results$`sub_df1[[1]]` <- c("Yes, to a major extent","Yes, to some extent","No, job on campus","No, job not on campus")

g <-ggplot(results, aes(`sub_df1[[1]]`, Percent, label=round(Valid_Percent,2), fill = Valid_Percent)) + geom_bar(stat='identity', fill="darkseagreen")+theme(plot.title = element_text(hjust = 0.5)) +geom_text(vjust=0)+labs(title = "Due to job, spent less time on campus than would like",x="Responses")

g

colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling





```

# Average hours per week spent:

```{r e5 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(81:85)

no_na <- sub_df1 %>% na.omit

results <- data.frame("Average_hours_per_week" = c("Doing research on campus","Preparing for class on campus","Travelling to and from university","Participating in extra-curricular activities","Providing care for dependents"), "N" = colSums(!is.na(sub_df1)), "Minimum" = c(min(no_na$Qe5_a),min(no_na$Qe5_b),min(no_na$Qe5_c),min(no_na$Qe5_d),min(no_na$Qe5_e)), "Maximum" = c(max(no_na$Qe5_a),max(no_na$Qe5_b),max(no_na$Qe5_c),max(no_na$Qe5_d),max(no_na$Qe5_e)) ,"Mean" = round(c(mean(no_na$Qe5_a),mean(no_na$Qe5_b),mean(no_na$Qe5_c),mean(no_na$Qe5_d),mean(no_na$Qe5_e)),3))


row.names(results)<-NULL

results  %>%
  kable() %>%
  kable_styling

```

# Age

```{r age2(table column 96) , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(96)

results <- get_summary(sub_df1)

g <-ggplot(results, aes(`sub_df1[[1]]`, Percent, label=round(Valid_Percent,2), fill = Valid_Percent)) + geom_bar(stat='identity', fill="darkseagreen") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0), plot.title = element_text(hjust = 0.54))+labs(title="Age Distribution",x="Age")

g

colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling

```

# Born in Canada

```{r f11 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(87)

results <- get_summary(sub_df1)

results$`sub_df1[[1]]` <- c("Yes","No")

colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling

```

# Gender

```{r GENDER (Column 92 in data table) , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(92)

results <- get_summary(sub_df1)

results$`sub_df1[[1]]` <- c("Male","Female")

g <- ggplot(results, aes("",Percent, fill=`sub_df1[[1]]`))+
  geom_bar(stat="identity", width =1, color="white") +
  coord_polar("y",start=0)+theme_void()+
  geom_text(aes(label = paste(round(Percent / sum(Percent) * 100, 1), "%")),
            position = position_stack(vjust = 0.5)) +
    labs(fill = "Gender",
       x = NULL,
       y = NULL,
       title = "Gender distribution") + theme(plot.title = element_text(hjust = 0.54))

g


colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling

```

# Ethnic/Racial/Cultural background

```{r e8/E8RECODE , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(93)

results <- get_summary(sub_df1)

results$`sub_df1[[1]]` <- c("White","Visible Minority","Other")

g <- ggplot(results, aes("",Percent, fill=`sub_df1[[1]]`))+
  geom_bar(stat="identity", width =1, color="white") +
  coord_polar("y",start=0)+theme_void()+
  geom_text(aes(label = paste(round(Percent / sum(Percent) * 100, 1), "%")),
            position = position_stack(vjust = 0.5)) +
    labs(fill = "Ethnic/Racial/Cultural background",
       x = NULL,
       y = NULL)

g

colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling

```

# Language spoke most often at home

```{r e9/Qe8_9@e9 , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(89)

results <- get_summary(sub_df1)

colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling
```

# Highest level of education completed by parents/guardians

```{r par_edu , include=TRUE, echo=FALSE , warning = FALSE}

sub_df1 <-  IT_rf2 %>% select(90)

results <- get_summary(sub_df1)

results$`sub_df1[[1]]` <- c("Did not finish high school","Graduated from high school","Some or completed college or CEGEP","Attended university without earning
degree","Completed Bachelor's Degree","Completed Master's Degree","Completed a Doctoral Degree","Completed a Professional Degree","Not sure")

g <-ggplot(results, aes(`sub_df1[[1]]`, Percent, label=round(Valid_Percent,2), fill = Valid_Percent)) + geom_bar(stat='identity', fill="darkseagreen") +geom_text(vjust=1)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0), plot.title = element_text(hjust = 0.54))+labs(title="Highest level of education completed by parents/guardians",x="Responses")

g

colnames(results)[1] <- " "

results  %>%
  kable() %>%
  kable_styling
```
